---
title: "Large PSM clustering"
author: "Paul DW Kirk"
date: "02/05/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Large PSM clustering

In order to get this Rmd file to run, I first changed the maximum memory available to R by following  the top answer given at: https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached

It would be worth checking to see if this (i.e. editing .Renviron) would fix the issue being encountered when using the maxpear function.  


```{r generatePSM}
rm(list = ls())

generatePSM <- function(n, seed = 1)
{
  nSamplesCluster1 <- round(0.1*n)
  nSamplesCluster2 <- round(0.2*n)
  nSamplesCluster3 <- round(0.3*n)
  nSamplesCluster4 <- n - (nSamplesCluster1 + nSamplesCluster2 + nSamplesCluster3)
  
  clusters <- c(rep(1,nSamplesCluster1), rep(2,nSamplesCluster2), 
                rep(3,nSamplesCluster3), rep(4,nSamplesCluster4))
  
  pilotPSM <- matrix(0, nrow = n, ncol = n)
  pilotPSM[1:nSamplesCluster1, 1:nSamplesCluster1]                   <- 1
  pilotPSM[(nSamplesCluster1+1):(nSamplesCluster1+nSamplesCluster2),
           (nSamplesCluster1+1):(nSamplesCluster1+nSamplesCluster2)] <- 1
  pilotPSM[(nSamplesCluster1+nSamplesCluster2+1):
             (nSamplesCluster1+nSamplesCluster2+nSamplesCluster3),
           (nSamplesCluster1+nSamplesCluster2+1):
             (nSamplesCluster1+nSamplesCluster2+nSamplesCluster3)]   <- 1
  pilotPSM[(nSamplesCluster1+nSamplesCluster2+nSamplesCluster3+1):n,
           (nSamplesCluster1+nSamplesCluster2+nSamplesCluster3+1):n] <- 1
  
  set.seed(seed)
  noiseMatrix <- matrix(rbeta(n*n,0.1,2), nrow = n, ncol = n)
  
  pilotPSM    <- abs(pilotPSM - noiseMatrix)
  newPSM      <- (pilotPSM + t(pilotPSM))/2
  return(list(psm = newPSM, clusters = clusters))
}

n <- 1000
psmResults   <- generatePSM(n)
trueClusters <- psmResults$clusters
psm1000      <- psmResults$psm


#Make symmetric
graphics.off()
pheatmap::pheatmap(psm1000, cluster_rows = F, cluster_cols = F)
hist(psm1000)
```

## Illustrate the use of HDBSCAN

We check to see how we can use HDBSCAN to perform clustering on the basis of the PSM.
```{r hdbscan}
library("ggplot2")
library("dbscan")

myDist         <- as.dist(1 - psm1000)

hdbscanResults <- hdbscan(myDist, minPts = round(0.05*n))

plot(hdbscanResults) # A summary dendrogram, showing the relationships between the clusters

# Generate annotations for rows and columns
annotation_col = data.frame(
  Cluster = factor(hdbscanResults$cluster+1L)
)

rownames(annotation_col) <- rownames(psm1000) <- 
  colnames(psm1000) <- paste0("V", seq(1,n))

pheatmap::pheatmap(psm1000, cluster_rows = F, cluster_cols = F, 
                   annotation_col = annotation_col, 
                   show_rownames = F, 
                   show_colnames = F)

print(mcclust::arandi(trueClusters, hdbscanResults$cluster))

```
## Assess how HDBSCAN scales with increasing n

```{r largePSMclustering}
library(tictoc)

#rangeOfSampleSizes <- c(100, 500, 1000, 1500, 2500, 5000, 10000, 20000, 
#                        50000, 1e5, 2e5, 5e5, 1e6, 5e6, 1e7, 1e8)

rangeOfSampleSizes <- c(100, 500, 1000, 1500, 2500, 5000, 10000, 27000)

nRepetitions  <- 5

timingsMatrix <- ariMatrix <- matrix(nrow = length(rangeOfSampleSizes), 
                                     ncol = nRepetitions) 

for(i in 1:length(rangeOfSampleSizes))
{
  print(paste("i = ",  i))
  for(j in 1:nRepetitions)
  {
    print(paste("j = ",  j))
    
    n                  <- rangeOfSampleSizes[i]
    print(n)
    psmResults         <- generatePSM(n)
    trueClusters       <- psmResults$clusters
    psm                <- psmResults$psm
    
    myDist             <- as.dist(1 - psm)
    
    tic()
    hdbscanResults     <- hdbscan(myDist, minPts = round(0.05*n))
    timed              <- toc(quiet = T)
    timingsMatrix[i,j] <- timed$toc - timed$tic
    
    ariMatrix[i,j]     <- mcclust::arandi(trueClusters, hdbscanResults$cluster)

    print(timingsMatrix[i,j])    
    
    
  }
  
  
}

timingsToPlot <- data.frame(n = rangeOfSampleSizes, meanTimes = rowMeans(timingsMatrix), lower = apply(timingsMatrix, 1, function(x) min(x)), upper = apply(timingsMatrix, 1, function(x) max(x)))

arisToPlot <- data.frame(n = rangeOfSampleSizes, meanARIs = rowMeans(ariMatrix), lower = apply(ariMatrix, 1, function(x) min(x)), upper = apply(ariMatrix, 1, function(x) max(x)))


pTimings <- ggplot(timingsToPlot, aes(n, meanTimes)) + geom_line() + geom_errorbar(aes(ymin = lower, ymax = upper), width = 2) 

pARIs <- ggplot(arisToPlot, aes(n, meanARIs)) + geom_line() + geom_errorbar(aes(ymin = lower, ymax = upper), width = 2) 


plot(pTimings)

plot(pARIs)


```

